### 2014-12-14 text/Chapter05.tex
Proving things (Chapter 5)

\begin{definition}
\label{def:structural-scalar}
A \emph{\cclrd{structural coeffect scalar}} $(\C, \cseq, \cpar, \cunit, \czero, \cleq)$ is a set 
$\C$ together with elements $\cunit, \czero \in \C$, relation $\cleq$ and binary operations 
$\cseq, \cpar$ such that $(\C, \cseq, \cunit)$ and $(\C, \cpar, \czero)$ are monoids and
$(\C, \cleq)$ is a pre-order. That is, for all $r,s,t\in \C$:
%
\begin{equation*}
\begin{array}{ccr}
r \;\cseq\; (s \;\cseq\; t) = (r \;\cseq\; s) \;\cseq\; t  &
\cunit \;\cseq\; r = r = r \;\cseq\; \cunit &
\textnormal{(monoid)}   
\\
r \;\cpar\; (s \;\cpar\; t) = (r \;\cpar\; s) \;\cpar\; t &
\czero \;\cpar\; r = r = r \;\cpar\; \czero &
\textnormal{(monoid)}   
\\
\textnormal{if}~~r\; \cleq\; s ~~\textnormal{and}~~s\; \cleq\; t~~\textnormal{then}~~r\; \cleq\; t&
t\; \cleq\; t &
\textnormal{(pre-order)}   
\end{array}
\end{equation*}
%
In addition, the following distributivity axioms hold:
\begin{align*}
\quad (\cclrd{r}\, \cpar\, \cclrd{s}) \;\cseq\; \cclrd{t} & = (\cclrd{r} \,\cseq\, \cclrd{t}) \;\cpar\; (\cclrd{s}\, \cseq\, \cclrd{t}) \\
\quad \cclrd{t} \;\cseq\; (\cclrd{r}\, \cpar\, \cclrd{s}) & = (\cclrd{t} \,\cseq\, \cclrd{r}) \;\cpar\; (\cclrd{t}\, \cseq\, \cclrd{s})
\end{align*}
\end{definition}

@@

@@

More generally, we could see the the coeffect annotations as a \emph{container} 
\cite{types-containers} that supports certain operations. This approach is used in 
Chapter~\ref{ch:unified} as a way of unifying the flat and structural systems.

@@

\subsection{Holes and substitution lemma}

@@

\begin{lemma}[Multi-nary substitution]
\label{thm:structural-substitution}
Given an expression with multiple holes filled by variables $x_i\!:\!\tau_i$ with coeffects $\cclrd{s_k}$:
%
\begin{equation*}
\coctx{\Gamma}{\aclrd{\textbf{r}}}~[\,\coctx{x_1\!:\!\tau_1}{\alift{\cclrd{s_1}}} \,|\, \ldots \,|\,
  \coctx{x_k\!:\!\tau_k}{\alift{\cclrd{s_k}}}\,] \vdash e_r : \tau_r
\end{equation*}
%
and a expressions $e_i$ with free-variable contexts $\Gamma_i$ annotated with $\aclrd{\mathbf{t_i}}$:
%
\begin{equation*}
\coctx{\Gamma_1}{\aclrd{\textbf{t}_1}} \vdash e_1 : \tau_1 
\quad \ldots \quad
\coctx{\Gamma_k}{\aclrd{\textbf{t}_k}} \vdash e_k : \tau_k
\end{equation*}
%
then substituting the expressions $e_i$ for variables $x_i$ results in an expression with a context
where the original holes are filled by contexts $\Gamma_i$ with coeffects $\cclrd{s_i} \aseq \aclrd{\mathbf{t_i}}$:
%
\begin{equation*}
\coctx{\Gamma}{\aclrd{\textbf{r}}}~[\,\coctx{\Gamma_s}{\cclrd{s_1}\aseq\,\aclrd{\textbf{t}_1}} \,|\, \ldots \,|\, 
  \coctx{\Gamma_s}{\cclrd{s_k}\aseq\,\aclrd{\textbf{t}_k}}\,] \vdash \subst{\subst{e_r}{x_1}{e_1}\ldots}{x_k}{e_k} : \tau_r
\end{equation*}
\end{lemma}
\begin{proof}
By induction over $\vdash$, using the multi-nary aspect of the substitution 
in the proof of the contraction case (see Appendix~\ref{sec:appendix-struct-cbn}).
\end{proof}

>>>>

\begin{definition}
\label{def:structural-scalar}
A \emph{\cclrd{structural coeffect scalar}} $(\C, \cseq, \cpar, \cunit, \czero, \cleq)$ is a set 
$\C$ together with elements $\cunit, \czero \in \C$, relation $\cleq$ and binary operations 
$\cseq, \cpar$ such that $(\C, \cseq, \cunit)$ and $(\C, \cpar, \czero)$ are monoids and
$(\C, \cleq)$ is a pre-order. That is, for all $r,s,t\in \C$:
%
\begin{equation*}
\begin{array}{ccr}
r \;\cseq\; (s \;\cseq\; t) = (r \;\cseq\; s) \;\cseq\; t  &
\cunit \;\cseq\; r = r = r \;\cseq\; \cunit &
\textnormal{(monoid)}   
\\
r \;\cpar\; (s \;\cpar\; t) = (r \;\cpar\; s) \;\cpar\; t &
\czero \;\cpar\; r = r = r \;\cpar\; \czero &
\textnormal{(monoid)}   
\\
\textnormal{if}~~r\; \cleq\; s ~~\textnormal{and}~~s\; \cleq\; t~~\textnormal{then}~~r\; \cleq\; t&
t\; \cleq\; t &
\textnormal{(pre-order)}   
\\
\end{array}
\end{equation*}
%
The following distributivity axioms hold:
\begin{align*}
\quad (\cclrd{r}\, \cpar\, \cclrd{s}) \;\cseq\; \cclrd{t} & = (\cclrd{r} \,\cseq\, \cclrd{t}) \;\cpar\; (\cclrd{s}\, \cseq\, \cclrd{t}) \\
\quad \cclrd{t} \;\cseq\; (\cclrd{r}\, \cpar\, \cclrd{s}) & = (\cclrd{t} \,\cseq\, \cclrd{r}) \;\cpar\; (\cclrd{t}\, \cseq\, \cclrd{s})
\end{align*}
%
In addition, we require the following two properties of $\czero, \cpar$ and $\cleq$:
\begin{equation*}
\begin{array}{l}
 \quad \cclrd{r}\,\cleq\,\cclrd{r'} \; \Rightarrow \; \forall\cclrd{s}.\;(\cclrd{r}\,\cseq\,\cclrd{s})\;\cleq\;(\cclrd{r'}\,\cseq\,\cclrd{s}) \\
 \quad \czero\;\cleq\;(\czero\,\cseq\,\cclrd{r})
\end{array}
\end{equation*}
\end{definition}

@@

Finally, we now also require two properties that relate the ordering $\cleq$ with coeffects composed
via $\cpar$ and with $\czero$. These were previously required for bottom-pointed substitution lemma 
(Lemma~\ref{thm:cbn-substitution-bot}). Here, we require them for all structural coeffects as it holds
for all our examples; it is needed for the structural version of the subtitution lemma 
(Lemma~\ref{thm:structural-substitution}).

@@

More generally, we could see the the coeffect annotations as \emph{containers} 
\cite{types-containers}. This approach is used in Chapter~\ref{ch:unified} to unify flat 
and structural systems.

@@

\subsection{Holes and substitution lemma}
\label{sec:struct-syntactic-subst}

@@

\begin{lemma}[Multi-nary substitution]
\label{thm:structural-substitution}
Given an expression with multiple holes filled by variables $x_i\!:\!\tau_i$ with coeffects $\cclrd{s_k}$:
%
\begin{equation*}
\coctx{\Gamma}{\aclrd{\textbf{r}}}~[\,\coctx{x_1\!:\!\tau_1}{\alift{\cclrd{s_1}}} \,|\, \ldots \,|\,
  \coctx{x_k\!:\!\tau_k}{\alift{\cclrd{s_k}}}\,] \vdash e_r : \tau_r
\end{equation*}
%
and a expressions $e_i$ with free-variable contexts $\Gamma_i$ annotated with $\aclrd{\mathbf{t_i}}$:
%
\begin{equation*}
\coctx{\Gamma_1}{\aclrd{\textbf{t}_1}} \vdash e_1 : \tau_1 
\quad \ldots \quad
\coctx{\Gamma_k}{\aclrd{\textbf{t}_k}} \vdash e_k : \tau_k
\end{equation*}
%
then substituting the expressions $e_i$ for variables $x_i$ results in an expression with a context
where the original holes are filled by contexts $\Gamma_i$ with coeffects $\cclrd{s_i} \aseq \aclrd{\mathbf{t_i}}$:
%
\begin{equation*}
\coctx{\Gamma}{\aclrd{\textbf{r}}}~[\,\coctx{\Gamma_1}{\cclrd{s_1}\aseq\,\aclrd{\textbf{t}_1}} \,|\, \ldots \,|\, 
  \coctx{\Gamma_k}{\cclrd{s_k}\aseq\,\aclrd{\textbf{t}_k}}\,] \vdash \subst{\subst{e_r}{x_1}{e_1}\ldots}{x_k}{e_k} : \tau_r
\end{equation*}
\end{lemma}
\begin{proof}
By induction over $\vdash$, using the multi-nary aspect of the substitution 
in the proof of the contraction case (see Appendix~\ref{sec:appendix-struct-cbn}).
\end{proof}
